---
title: "MSJemutai_PS_3"
author: "Mercy Salome Jemutai"
date: "9/15/2019"
output:
  word_document: default
  html_document: default
---


#https://www.zevross.com/blog/2014/08/04/beautiful-plotting-in-r-a-ggplot2-cheatsheet-3/#TOC

##QN 1 : Identify the type of variables in this data set (numeric, categorical) and 
##Display the most appropriate graph based on variable type using basic graphical functions.

#read in data, and attach 


```{r}
Telomeres<-read.csv("C:\\Users\\mjemutai\\Downloads\\TELOMERES.csv")
```

```{r}
attach(Telomeres)
```

```{r}
#use the library()command

library(ggplot2)

#using head() and tail commands to  see first 6 and last six rows of the dataframe (if large data set)
head(Telomeres)

tail(Telomeres)

# On running the above, we can tell that the variables we have, years and telomere.length, are both numerical


``` 

#Display using geom_point since we using a scatterplot to check the relationship between the two variables


```{r}

#https://www.rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf
#https://www.zevross.com/blog/2014/08/04/beautiful-plotting-in-r-a-ggplot2-cheatsheet-3/#add-a-title-ggtitle-or-labs
#FORMAT:data=data.frame path)+geom_function(mapping=aes(MAPPINGS))

#gplot
#add color, size, title, background color

#
ggplot(data = Telomeres)+geom_point(mapping = aes(x = years,y=telomere.length), color="orange",size=4)+ggtitle("Relationship between length of telomere and years")+theme(panel.background = element_rect(fill = 'purple'))



       
```


```{r}

#smooth to view trend
ggplot(data = Telomeres)+ geom_smooth(mapping =aes(x=years,y=telomere.length))





    
   #From the trend, overall, as time(years) increase, the length of the telomere decreases!

#it is likely that the telomeres of parents/caregivers of very sick children  shorten as a result of chronic stress, perhaps, as they spend more time looking after sick children

```



#QN 2:  Improve the scatterplot you produced from the hippocampus.csv dataset in Set 1

#read in data, and attach


```{r}


hippocampus<-read.csv("C:\\Users\\mjemutai\\Downloads\\hippocampus_218P.csv")
```



```{r}
attach(hippocampus)
```


```{r}
head(hippocampus)

tail(hippocampus)

#we don't have to check this though!
```
```{r}
#Let's plot using geom_point()

#add title, color, shape, background, and size of shape

ggplot(data = hippocampus, mapping = aes(x=posterior_hippo_vol, y=months_as_taxi_driver))+geom_point(shape=11, size=3, color="red")+ggtitle("Hippocampus data")+theme(panel.background = element_rect(fill = 'light green'))
                                                                                                  
                                                                                                





```

#QN 3:  Create a graph depicting the frequency distribution of the 35 mass measurements




```{r}

#Input data as a vector using c() function
mass<-c(0.047, 0.037, 0.041, 0.045, 0.039, 0.064, 0.064, 0.065, 0.079, 0.070, 0.066,
0.059, 0.075, 0.079, 0.090, 0.069, 0.066, 0.078, 0.066, 0.066, 0.055, 0.046,
0.056, 0.067, 0.075, 0.048, 0.077, 0.081, 0.066, 0.172, 0.080, 0.078, 0.048,
0.096, 0.097)




```



```{r}

#change the data from vector to a data frame so we can use ggplot2

mass2<-as.data.frame(mass)

```




```{r}

#histograms use *geom_histogram*, boxplots use *geom_boxplot*

#we can use histograms to represent frequency distribution of numeric data

#https://www.datacamp.com/community/tutorials/make-histogram-ggplot2
#ggplot(data = mass2 )+geom_histogram(aes(mass2$mass)) - simple histogram


ggplot(data = mass2,aes(mass2$mass))+geom_histogram(binwidth = 0.01,col="black",fill="orange") + labs(title="Histogram for mass", x="mass",y="Count")
                                                    
                                                    
  #change width of the histogram bins - the error messsage!                                            
                                              



      


```



#Describe the shape of the frequency distribution. What are its main features? (Hint: are there any outliers?).



```{r}
#The histogram is somehow bell-shaped although increasing the width of the bins makes it  positively skewed. 


#Yes, we have an outlying data near mass = 0.153


```



```{r}
detach(Telomeres)
detach(hippocampus)

```



